Active Typechecking and Translation: A Safe Language-Internal Extension Mechanism

### 2013-10-12 esop14/att-esop14.tex
i removed a couple of sentences that are just circular
it'd be nice to have a better way to lead into the next section, but i
think the point is at least more clearly made this way.

Statically-typed languages typically make a distinction between \emph{expressions}, which describe run-time computations, and type-level constructs like types, type aliases and datatype declarations. The design described above suggests we may now need to add another layer to our language, an {extension language}, where extensions can be declared and implemented. In fact, we will show that \textbf{the most natural place for type system extensions is within the type-level language}. The intuition is that extensions to a language's static semantics will need to manipulate types as values at compile-time. Many languages already allow users to write type-level functions for various reasons, effectively supporting this notion of types as values at compile-time (see Sec. \ref{related-work} for examples). The type-level language is often constrained by its own type system (where the types of type-level values are called \emph{kinds} for clarity) that prevents type-level functions from causing problems during compilation. This is precisely the structure that a distinct extension layer would have, and so we will demonstrate that it is quite natural to unify the two in this work.

>>>>

Statically-typed languages typically make a distinction between \emph{expressions}, which describe run-time computations, and type-level constructs like types, type aliases and datatype declarations. The design described above suggests we may now need to add another layer to our language, an {extension language}, where extensions can be declared and implemented. In fact, we will show that the most natural place for type system extensions is within the type-level language, with the intuition being that extensions to a language's static semantics will need to manipulate types as values at compile-time.

% Many languages already allow users to write type-level functions, effectively supporting this notion of types as values at compile-time (see Sec. \ref{related-work} for examples). The type-level language is often constrained by its own type system (where the types of type-level values are called \emph{kinds} for clarity) that prevents type-level functions from causing problems during compilation. This is precisely the structure that a distinct extension layer would have, and so we will demonstrate that it is quite natural to unify the two in this work.
%
% ^ the rest of this paragraph is a bit repetitive and circular. you wish to make the point that the type-level is the best place to include type system extensions because you have types as values at compile time. feel free to put it back, it just seems clearer to me this way.
